Canopy leaf area of a mature evergreen *Eucalyptus* woodland does not respond to elevated atmospheric [CO~2~] but tracks water availability

### 2015-05-21 manuscript.Rmd
Fixed PET estimate from Zomer; moved R code in manuscript to manuscript_calcs.R

```{r echo=FALSE, message=FALSE}
library(knitr)

load("cache/lai_workspace.RData")
source("load.R")

source("R/functions-figures.R")
source("R/figures.R")
source("R/derivSimulCI.R")

library(broom)
library(lme4)
library(lmerTest)
library(car)
library(reporttools)

# shortcut to formatPval
pval <- function(...)formatPval(..., includeEquality=TRUE)

# Number of digits in printing of numbers
# options(scipen = 1, digits = 3)

# Shortcut to significant digits formatting
f <- function(x, digits=2)sprintf(x, fmt=paste0("%.",digits,"f"))

# Some definitirons of constants
.gambase <- 18
.maxSD <- 0.03

# Main results
df <- facegap_cloudy_byring

# Some pre-calculations
caliblm <- calibrateToDrought(df)$lmfit

# Number of months of study
r <- range(df$Date)
r[2] <- maxdate
nummon <- as.numeric(floor(difftime(r[2],r[1],units="weeks")/(52/12)))

# Range mean LAI
dfa <- summaryBy(LAI ~ Date, FUN=mean, na.rm=TRUE, data=df)
rangeLAI <- round(range(dfa$LAI),1)

# LAI ~ BA
balm <- lm(LAI.mean ~ BA, data=ba)

# Comparison flat canopy photo with transmittance
lmphoto <- lm(LAI.PAR.mean ~ LAI, data=flatcan_byring)

# Significance CO2 on LAI anomaly
df$Time <- as.factor(df$Date - min(df$Date))
m3 <- lmer(LAIanomaly ~ treatment*Time + (1|Ring), data=df)
LanomAOV <- Anova(m3)

# Average CI width in Fig 4.
fit <- smoothplot(Date, LAIanomaly, g=treatment, R="Ring", 
                  data=facegap_cloudy_byring, plot=FALSE, kgam=.gambase)
pa <- predict(fit$ambient, data.frame(X=as.numeric(facegap_cloudy_byring$Date)), se.fit=TRUE)
pe <- predict(fit$elevated, data.frame(X=as.numeric(facegap_cloudy_byring$Date)), se.fit=TRUE)
LanomCIamb <- mean(pa$se.fit)*2
LanomCIele <- mean(pe$se.fit)*2

# Detectable Ca effect on LAI
Ldetect <- 100 * (LanomCIamb + LanomCIele) / mean(ba$LAI.mean)

# Litter - dLAI correlations Figure 4
dl <- dLAIlitter
dl$Y <- with(dl, dLAI * 30.5/ndays)
dl$X <- with(dl, dLAIlitter.mean * 30.5/ndays)
dlup <- dl[dl$LAIchange == "increasing",]
dldown <- dl[dl$LAIchange == "decreasing",]
lmlitup <- lm(Y ~ X, data=dlup)
lmlitdown <- lm(Y ~ X, data=dldown)

# Leaf lifespan
LLtest <- tidy(t.test(LL ~ treatment,data=ba))

# Canopy photo - Ltau comparison
lmcanphot <- lm(LAI.PAR.mean ~ LAI, data=flatcan_byring)

# With optimized k
flatcan_byring2 <- agg_flatcan(flatcan, by="Ring", k=kopt)
flatcan_byring2 <- add_PARLAI_to_flatcan(facegap_cloudy_byring,flatcan_byring2)
lmcanphot2 <- lm(LAI.PAR.mean ~ LAI, data=flatcan_byring2)

```

@@

\
We estimated *L* from near-continuous measurements of canopy transmittance in a *Eucalyptus*-dominated mature woodland exposed to eC~a~, the EucFACE experiment. The woodland at the EucFACE is frequently water-limited, as judged from the deficit between potential evapotranspiration (long-term mean of 1500mm; BOM) and mean annual precipitation of 800mm. The strong seasonal and interannual variability in precipitation at the site allowed us to quantify the dynamics of *L*, and its components (litter fall and production) in response to water availability. We hypothesized that eC~a~ would result in an increase in *L*, particularly during periods of water limitation. In addition, we expected that the low *L* that is typical of native *Eucalyptus*-dominated woodlands [@Whitehead2004] allowed a positive response of *L* to eC~a~ [cf. @Norby2011], either via an increase in the rate of leaf production or leaf retention (leaf lifespan). Our second objective was to quantify the temporal dynamics of *L*, in particular in relation to fluctuating water availability.

@@

## Site description and experimental design
The EucFACE experiment is located in a remnant patch  of native Cumberland Plain woodland [@Tozer2003] on an ancient alluvial floodplain in western Sydney (Australia, 33^o^ 37'S, 150^o^ 44'E, 30 m a.s.l.). The vegetation consists of a native woodland undisturbed for ca. 70 years and maintained in this location since the 1880’s. The site is characterized by a humid temperate-subtropical transitional climate with a mean annual temperature of 17 ^o^C and mean annual precipitation (P) of 800 mm (Bureau of Meteorology, station 067105 in Richmond, NSW Australia; http://www.bom.gov.au). The potential evapotranspiration (PET) is ca. 1500mm (*to be refined!*), giving a ratio of P/PET of ca. 0.53. Over the last 30 years, monthly PET has exceeded monthly P 78% of the time, indicating very frequent water limitation.

@@

## Leaf litter production
In each ring, 8 litter cloth baskets with an area of 0.1979 m^2^ were installed at 0.5m above the forest floor. Litter was collected monthly since October 2012, and leaves were separated from other components such as twigs, bark, seeds, insect frass, etc. Litter production was converted to units of *L* using an estimate of specific leaf area (SLA) determined from 329 leaves collected from the tree canopies from May 2012 to May 2014 (`r get_sla()` cm^2^ g^-1^, SE=2.14). We did not find a significant effect of C~a~ treatment on SLA (Ellsworth, unpublished data; Ben Moore, unpublished data).

>>>>

```{r echo=FALSE, message=FALSE}
source("manuscript_calcs.R")
```

@@

\
We estimated *L* from near-continuous measurements of canopy transmittance in a *Eucalyptus*-dominated mature woodland exposed to eC~a~, the EucFACE experiment. The woodland at the EucFACE is frequently water-limited, as judged from the deficit between potential evapotranspiration [long-term mean of 1350mm; @Zomer2008] and mean annual precipitation of 800mm. The strong seasonal and interannual variability in precipitation at the site allowed us to quantify the dynamics of *L*, and its components (litter fall and production) in response to water availability. We hypothesized that eC~a~ would result in an increase in *L*, particularly during periods of water limitation. In addition, we expected that the low *L* that is typical of native *Eucalyptus*-dominated woodlands [@Whitehead2004] allowed a positive response of *L* to eC~a~ [cf. @Norby2011], either via an increase in the rate of leaf production or leaf retention (leaf lifespan). Our second objective was to quantify the temporal dynamics of *L*, in particular in relation to fluctuating water availability.

@@

## Site description and experimental design
The EucFACE experiment is located in a remnant patch  of native Cumberland Plain woodland [@Tozer2003] on an ancient alluvial floodplain in western Sydney (Australia, 33^o^ 37'S, 150^o^ 44'E, 30 m a.s.l.). The vegetation consists of a native woodland undisturbed for ca. 70 years and maintained in this location since the 1880’s. The site is characterized by a humid temperate-subtropical transitional climate with a mean annual temperature of 17 ^o^C and mean annual precipitation (P) of 800 mm (Bureau of Meteorology, station 067105 in Richmond, NSW Australia; http://www.bom.gov.au). An estimate of potential evapotranspiration (PET) at the site (1350mm) was from the Global-PET database [http://www.cgiar-csi.org, @Zomer2008], giving a ratio of P/PET of ca. 0.59. In addition, over the last 30 years, monthly PET has exceeded monthly P 78% of the time, indicating very frequent water limitation.

@@

## Leaf litter production
In each ring, 8 litter cloth baskets with an area of 0.1979 m^2^ were installed at 0.5m above the forest floor. Litter was collected monthly since October 2012, and leaves were separated from other components such as twigs, bark, seeds, insect frass, etc. Litter production was converted to units of *L* using an estimate of specific leaf area (SLA) determined from 329 leaves collected from the tree canopies from May 2012 to May 2014 (`r get_sla()` cm^2^ g^-1^, SE=2.14). We did not find a significant effect of C~a~ treatment on SLA (Ellsworth, unpublished data) at any of the collection dates.
